### 2024-01-05 experiments/notebooks/figure2.qmd

using MLJBase
niter = 1000
jem = mach.model.model.jem
_batch_size = mach.model.model.batch_size
X = Float32.(MLJBase.matrix(X))
if typeof(jem.sampler) <: ConditionalSampler

    plts = []
    for target in 1:2
        X̂ = generate_conditional_samples(jem, _batch_size, target; niter=niter) 
        ex = extrema(hcat(X,X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=100)
        x2 = range(1.0f0.*ylims...,length=100)
        plt = Plots.contour(
            x1, x2, (x, y) -> softmax(jem([x, y]))[target], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(X[1,:], X[2,:], color=Int.(labels.refs), group=Int.(labels.refs), alpha=0.5)
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target], size(X̂,2)), 
            group=repeat([target], size(X̂,2)), 
            shape=:star5, ms=10
        )
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400))
    display(plt)
end
savefig(plt, joinpath(output_images_path, "poc_conditional_samples.png"))
```

@@

>>>>

using MLJBase
niter = 1000
jem = mach.model.model.jem
_batch_size = mach.model.model.batch_size
X = Float32.(MLJBase.matrix(X))
gen_samples = Dict{Any,Any}()
if typeof(jem.sampler) <: ConditionalSampler

    plts = []
    for target in 1:2
        X̂ = generate_conditional_samples(jem, _batch_size, target; niter=niter) 
        gen_samples[target] = X̂
        ex = extrema(hcat(X,X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=100)
        x2 = range(1.0f0.*ylims...,length=100)
        plt = Plots.contour(
            x1, x2, (x, y) -> softmax(jem([x, y]))[target], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(X[1,:], X[2,:], color=Int.(labels.refs), group=Int.(labels.refs), alpha=0.5)
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target], size(X̂,2)), 
            group=repeat([target], size(X̂,2)), 
            shape=:star5, ms=10
        )
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400))
    display(plt)
end
savefig(plt, joinpath(output_images_path, "poc_conditional_samples.png"))
```

@@

## Simple CE

```{julia}
Random.seed!(2024)
M = fit_model(counterfactual_data, :Linear)
gen = WachterGenerator(opt=Flux.Descent(0.001))
factual_label =  levels(labels)[2]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[1]
factual = predict_label(M, counterfactual_data, x_factual)[1]
ce = generate_counterfactual(
    x_factual, target, counterfactual_data, M, gen;
    converge_when=:generator_conditions,
)
plt = plot(
    ce, dpi=300, size=(500,500), 
    cbar=false, axis=nothing, legend=false,
    palette = col_pal,
)
savefig(plt, joinpath(output_images_path, "simple_ce.png"))
```

## True Density

```{julia}
using KernelDensity
using StatsPlots

target_cls_mask = counterfactual_data.output_encoder.labels .== 1
X_target = counterfactual_data.X[:,target_cls_mask]

x = X_target[1,:]
y = X_target[2,:]

k = kde((x, y))

plt = contourf(
    k,
	c = :BuPu,
	linewidth = 0,
    size=(500,500),
    dpi=300,
    colorbar=false,
    axis=nothing,
    legend=false,
    palette = col_pal,
)

scatter!(
    counterfactual_data, alpha=0.3,
    xlims = extrema(x),
    ylims = extrema(y),
)

ce_path = hcat(path(ce)...)
scatter!(
    ce_path[1,:], ce_path[2,:], 
    color=predict_label(M, counterfactual_data, ce_path), ms=10, alpha=0.5,
    label=""
)
savefig(plt, joinpath(output_images_path, "density_true.png"))
```

## Posterior

```{julia}
x = gen_samples[target][1,:]
y = gen_samples[target][2,:]

k = kde(
    (x, y); 
    bandwidth=(0.1,0.1),
    boundary=(extrema(X_target[1,:]),extrema(X_target[2,:]))
)

plt = contourf(
    k,
	c = :BuPu,
	linewidth = 0,
    size=(500,500),
    dpi=300,
    colorbar=false,
    axis=nothing,
    legend=false,
    palette = col_pal,
)

scatter!(
    counterfactual_data, alpha=0.3,
    xlims = extrema(X_target[1,:]),
    ylims = extrema(X_target[2,:]),
)

ce_path = hcat(path(ce)...)
scatter!(
    ce_path[1,:], ce_path[2,:], 
    color=predict_label(M, counterfactual_data, ce_path), ms=10, alpha=0.5,
    label=""
)

scatter!(
    plt, gen_samples[1][1,:], gen_samples[1][2,:],
    shape=:star5, ms=10, color=:yellow, alpha=0.5, label=""
)

savefig(plt, joinpath(output_images_path, "density_model.png"))
```
